---
title: "ARIMA"
output: html_notebook
---

```{r}
library(astsa)
library(xts)
library(tsbox)
library(MASS)
```


Must run functions in Detrend first



```{r}
dev.new(width=5, height=6)
acf(data,na.action=na.pass,main='ACF of detrended data')
pacf(data,na.action=na.pass,main='ACF of detrended data')
#plot(data,type='l')
```

1) ACF does not decay, try differencing
2) Both ACF and PACF show clear signs of seasonality present in correlation structure

```{r}
library(tseries)
library(imputeTS)
adf.test(na_seadec(diff(data)))

```
```{r}
acf(diff(data),na.action=na.pass,main='ACF of Δ(detrended data)')
pacf(diff(data),na.action=na.pass,main='PACF of Δ(detrended data)')
plot(diff(data),type='l',main='Plot of Δ(detrended data)',ylab='Δ(detrended data)')
```


1) Still need to difference to account for seasonality in ACF

```{r}
acf(diff(diff(data),lag=7),na.action=na.pass,main='ACF of Δ(Δ(detrended data),lag = 7))')
pacf(diff(diff(data),lag=7),na.action=na.pass,main='PACF of Δ(Δ(detrended data),lag = 7))')
plot(diff(diff(data),lag=7),type='l',main='Plot of Δ(Δ(detrended data),lag = 7))',ylab='data')
```

1)Both ACF, PACF displaying geometric decay. Now ready to fit AR/MA terms
2)Still appears to be yearly seasonality present in the fit data. May be worth exploring weekly/monthly frequencies to allow for fitting yearly trends. 


```{r}
##WARNING: This takes several minutes to execute. Only run if models need to be refit
#Ctrl-Shift-C to block comment/uncomment

AICtable = matrix(NA, 4, 4)
BICtable = matrix(NA, 4, 4)
dimnames(AICtable) =
list(paste("p =", 0:3), paste("q =", 0:3))

for (p in 0:3) {
for (q in 0:3) {
  #Had to use method='ML' since conditional sum of squares was returning nonstationary AR coefficients.
gt.arma=arima(diff(diff(data),lag=7), order=c(0,0,0), seasonal =
list(order = c(p,0,q), period = 7), include.mean
= FALSE)
AICtable[p+1, q+1] = AIC(gt.arma)
BICtable[p+1, q+1] = BIC(gt.arma)
}}
AICtable
BICtable
```

SARIMA(1,1,2)_7 appears to be the best choice based on BIC.

```{r}
candmod = arima(data, order=c(0,1,0), seasonal =
list(order = c(1,1,2), period = 7), include.mean
= FALSE)

plot(candmod$residuals,main = 'SARIMA Residuals')
acf(candmod$residuals,na.action = na.pass, main = 'ACF of SARIMA residuals')
pacf(candmod$residuals,na.action = na.pass, main = 'PACF of SARIMA residuals')
```
1)Improved seasonal effect in the PACF
2)Ready to fit ARIMA(p,1,q)

```{r}

##WARNING: This takes several minutes to execute. Only run if models need to be refit
#Ctrl-Shift-C to block comment/uncomment

AICtable = matrix(NA, 4, 4)
BICtable = matrix(NA, 4, 4)
dimnames(AICtable) =
list(paste("p =", 0:3), paste("q =", 0:3))

for (p in 0:3) {
for (q in 0:3) {
  #Had to use method='ML' since conditional sum of squares was returning nonstationary AR coefficients.
gt.arma=arima(data, order=c(p,1,q), seasonal =
list(order = c(1,1,2), period = 7), include.mean
= FALSE)
AICtable[p+1, q+1] = AIC(gt.arma)
BICtable[p+1, q+1] = BIC(gt.arma)
}}
AICtable
BICtable
```

Table suggests that ARIMA(1,1,3) is optimal.

```{r}
candmod = arima(data, order=c(1,1,3), seasonal =
list(order = c(1,1,2), period = 7), include.mean
= FALSE)

plot(candmod$residuals,main = 'ARIMA(1,1,3)xSARIMA(1,1,2)_7')
acf(candmod$residuals,na.action = na.pass,main = 'ARIMA(1,1,3)xSARIMA(1,1,2)_7')
pacf(candmod$residuals,na.action = na.pass,main = 'ARIMA(1,1,3)xSARIMA(1,1,2)_7')

```

```{r}
data = ts(logdata,freq=7
          )


data = na_seadec(data)
candmod = arima(data, order=c(1,1,3), seasonal =
list(order = c(1,1,2), period = 7), xreg=cbind(time(data),factor(cycle(data))), include.mean
= FALSE,method='CSS')

predtime = ts(frequency = 7,start = time(data)[3722]+1/7, end =  time(data)[3722]+4)
predtime=time(predtime)


ARIMAforecast = predict(candmod,n.ahead=28,newxreg = cbind(predtime,factor(cycle(predtime))))

postCOVIDdata = window(xtsdata, start = "2020-3-10", end = "2020-4-10")
postCOVIDtsdata = ts_ts(postCOVIDdata)
postCOVIDtsdata = ts(postCOVIDtsdata,frequency = 7,start = time(data)[3722], end =  time(data)[3722]+4)

plot(exp(data),xlim=c(510,540),ylim=c(10000000,25000000),main='Effect of COVID-19 on Steam Users')
lines(exp(ARIMAforecast$pred),col='2',lty=2)
lines(postCOVIDtsdata,lty=2)

```

```{r}
mean(postCOVIDtsdata[2:29] - exp(ARIMAforecast$pred))

```