---
title: "Transformations"
output: html_notebook
---


```{r}
library(astsa)
library(xts)
library(tsbox)
library(MASS)
```

```{r}
setwd('~/Documents/projects/MATH598_Final')
```
```{r}
userData = read.csv('/home/mholswade/Documents/projects/MATH598_Final/data/steamUsers.csv')
```

Data Cleaning
```{r}
times = as.Date(userData$DateTime,format='%Y-%m-%d')
xtsdata = xts(userData$Users,times)
tsdata = ts_ts(xtsdata)
```

```{r}
preCOVIDdata = window(xtsdata, start = "2010-1-1", end = "2020-3-10")
preCOVIDtsdata = ts_ts(preCOVIDdata)
```

Check histograms
```{r}
hist(preCOVIDtsdata,main='Histogram of pre COVID-19 user numbers',xlab='Number of Users')
plot(preCOVIDtsdata,main='Plot of pre COVID-19 User Numbers',ylab='Number of Users')
```

Data is roughlly uniform, will try some transformations for possible improvements to the fit



```{r}
logdata = log(preCOVIDtsdata)
hist(logdata,main='Plot of pre COVID-19 User Numbers (log transform)',xlab='Number of Users')
plot(logdata,main='Plot of pre COVID-19 User Numbers (log transform)',ylab='Number of Users')
```
change data to form with cycles
```{r}
data = ts(logdata,freq=7
          )

```

```{r}
model = lm(data ~ time(data) + factor(cycle(data)))
modelfit = ts(fitted(model),start=start(data),frequency=frequency(data))

plot(data,main='Linear regression',ylab='Number of Users')
lines(modelfit,col=2,lty=2)
plot(model$residuals,type='l',main='Residuals of Model',xlab='Time',ylab='Residual Value')
```

```{r}
time2 = time(data)^2
quadmodel = lm(data ~ time2 + time(data) + factor(cycle(data)))
quadmodelfit = ts(fitted(quadmodel),start=start(data),frequency=frequency(data))

plot(data)
lines(quadmodelfit,col=2,lty=2)
plot(quadmodel$residuals,type='l')
```
```{r}

simpmodel = lm(data ~ + factor(cycle(data)))
simpmodelfit = ts(fitted(simpmodel),start=start(data),frequency=frequency(data))

plot(data)
lines(simpmodelfit,col=2,lty=2)
plot(simpmodel$residuals,type='l')
```

```{r}

diffmodel = lm(diff(data) ~ + factor(cycle(data)))
simpmodelfit = ts(fitted(),start=start(data),frequency=frequency(data))

plot(data)
lines(simpmodelfit,col=2,lty=2)
plot(simpmodel$residuals,type='l')
```

```{r}
data = model$residuals

```

Conclusions:
1) 5 year data does not seem to be a long enough look to accurately establish an overall trend
2) When 10 year trend is looked at, linear and quad fits are nearly identical.
3) Transforming the data with candidate transformations did not seem to produce any meaningful results

