---
title: "ARIMA"
output: html_notebook
---

```{r}
library(astsa)
library(xts)
library(tsbox)
library(MASS)
```


Must run functions in Detrend first

```{r}
data = detrended

```

```{r}

acf(data,na.action=na.pass)
pacf(data,na.action=na.pass)
plot(data,type='l')
```

1) ACF does not decay, try differencing
2) Both ACF and PACF show clear signs of seasonality present in correlation structure

```{r}
acf(diff(data),na.action=na.pass)
pacf(diff(data),na.action=na.pass)
plot(diff(data),type='l')
```


1) Still need to difference to account for seasonality in ACF

```{r}
acf(diff(diff(data),lag=7),na.action=na.pass)
pacf(diff(diff(data),lag=7),na.action=na.pass)
plot(diff(diff(data),lag=7),type='l')
```

1)Both ACF, PACF displaying geometric decay. Now ready to fit AR/MA terms
2)Still appears to be yearly seasonality present in the fit data. May be worth exploring weekly/monthly frequencies to allow for fitting yearly trends. 


```{r}
##WARNING: This takes several minutes to execute. Only run if models need to be refit
#Ctrl-Shift-C to block comment/uncomment

# AICtable = matrix(NA, 5, 5)
# BICtable = matrix(NA, 5, 5)
# dimnames(AICtable) =
# list(paste("p =", 0:4), paste("q =", 0:4))
# 
# for (p in 0:4) {
# for (q in 0:4) {
#   #Had to use method='ML' since conditional sum of squares was returning nonstationary AR coefficients.
# gt.arma=arima(data, order=c(0,1,0), seasonal =
# list(order = c(p,1,q), period = 7), include.mean
# = FALSE,method="ML")
# AICtable[p+1, q+1] = AIC(gt.arma)
# BICtable[p+1, q+1] = BIC(gt.arma)
# }}
# AICtable
# BICtable
```

SARIMA(1,1,2)_7 appears to be the best choice based on BIC.

```{r}
candmod = arima(data, order=c(0,1,0), seasonal =
list(order = c(1,1,2), period = 7), include.mean
= FALSE)

plot(candmod$residuals)
acf(candmod$residuals,na.action = na.pass)
pacf(candmod$residuals,na.action = na.pass)
```
1)Improved seasonal effect in the PACF
2)Ready to fit ARIMA(p,1,q)

```{r}

##WARNING: This takes several minutes to execute. Only run if models need to be refit
#Ctrl-Shift-C to block comment/uncomment

AICtable = matrix(NA, 4, 4)
BICtable = matrix(NA, 4, 4)
dimnames(AICtable) =
list(paste("p =", 0:3), paste("q =", 0:3))

for (p in 0:3) {
for (q in 0:3) {
  #Had to use method='ML' since conditional sum of squares was returning nonstationary AR coefficients.
gt.arma=arima(data, order=c(p,1,q), seasonal =
list(order = c(1,1,2), period = 7), include.mean
= FALSE)
AICtable[p+1, q+1] = AIC(gt.arma)
BICtable[p+1, q+1] = BIC(gt.arma)
}}
AICtable
BICtable
```