---
title: "Transformations"
output: html_notebook
---


```{r}
library(astsa)
library(xts)
library(tsbox)
library(MASS)
```

```{r}
setwd('~/Documents/projects/MATH598_Final')
```

Data Cleaning
```{r}
userData = read.csv('/home/mholswade/Documents/projects/MATH598_Final/data/steamUsers.csv')
```


```{r}
times = as.Date(userData$DateTime,format='%Y-%m-%d')
xtsdata = xts(userData$Users,times)
tsdata = ts_ts(xtsdata)
```

```{r}
preCOVIDdata = window(xtsdata, start = "2010-1-1", end = "2020-3-10")
preCOVIDtsdata = ts_ts(preCOVIDdata)
```

Check histograms
```{r}
hist(preCOVIDtsdata)
plot(preCOVIDtsdata)
```

Data is roughlly uniform, will try some transformations for possible improvements to the fit



```{r}
logdata = log(preCOVIDtsdata)
hist(logdata)
plot(logdata)
```
change data to form with cycles
```{r}
data = ts(preCOVIDtsdata,freq=7
          )

```

```{r}
model = lm(data ~ time(data) + factor(cycle(data)))
modelfit = ts(fitted(model),start=start(data),frequency=frequency(data))

plot(data)
lines(modelfit,col=2,lty=2)
plot(model$residuals,type='l')
```

```{r}
time2 = time(data)^2
quadmodel = lm(data ~ time2 + time(data) + factor(cycle(data)))
quadmodelfit = ts(fitted(quadmodel),start=start(data),frequency=frequency(data))

plot(data)
lines(quadmodelfit,col=2,lty=2)
plot(quadmodel$residuals,type='l')
```
```{r}

simpmodel = lm(data ~ + factor(cycle(data)))
simpmodelfit = ts(fitted(simpmodel),start=start(data),frequency=frequency(data))

plot(data)
lines(simpmodelfit,col=2,lty=2)
plot(simpmodel$residuals,type='l')
```

```{r}

diffmodel = lm(diff(data) ~ + factor(cycle(data)))
simpmodelfit = ts(fitted(),start=start(data),frequency=frequency(data))

plot(data)
lines(simpmodelfit,col=2,lty=2)
plot(simpmodel$residuals,type='l')
```

```{r}
detrended = model$residuals

```

Conclusions:
1) 5 year data does not seem to be a long enough look to accurately establish an overall trend
2) When 10 year trend is looked at, linear and quad fits are nearly identical.
3) Transforming the data with candidate transformations did not seem to produce any meaningful results

